Solving the maximum independant set problem with tensor networks

### 2021-05-26 paper/paper.tex
finite field method

%\documentclass[a4paper,superscriptaddress,11pt]{quantumarticle}
%\documentclass[aps,twocolumn,longbibliography,english,superscriptaddress]{revtex4-1}
\documentclass{article}
\usepackage{iclr2021_conference}
%\documentclass[a4paper,superscriptaddress,11pt]{article}
\pdfoutput=1
\usepackage[colorlinks=true,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{tabularx}
\usepackage{quoting}
\usepackage{upquote}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage[framemethod=TikZ]{mdframed}
\usepackage{wrapfig}
%\usepackage{caption}
%\usepackage[plain]{algorithm}
\usepackage[ruled, vlined]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{rotating}
%\usepackage{cite}
\usepackage{booktabs}
%\usepackage{unicode-math}
%\usepackage{algorithm}% http://ctan.org/pkg/algorithm
%\usepackage{algpseudocode}% http://ctan.org/pkg/algpseudocode
\usepackage{xcolor}% http://ctan.org/pkg/xcolor
\makeatletter
\newsavebox{\@brx}
\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\makeatother
\usepackage{bbm}
\usepackage{jlcode}
\usepackage{graphicx}
\usepackage{amsmath,color,amsthm}
\usepackage{mathrsfs}
\usepackage{float}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{indentfirst}
\usepackage{txfonts}
\usepackage[epsilon, tsrm, altpo]{backnaur}

@@

\begin{abstract}
	Solving the maximum independent set size problem: the maximum independent set size, the independance polynomial and the optimal configuration.
\end{abstract}

@@

For larger graph, we hight recommend using two additional packages,
\texttt{TropicalGEMM} for BLAS speed tropical matrix multiplication and
\texttt{OMEinsumContractionOrders} for hyper optimized contraction order (the KaHyPar approach) ~\cite{Gray2021,Pan2021}.

@@

Then we contract this graph with einsum \texttt{a,b,c,d,ab,ad,bc,bd,cd,ce->}.
The result is a scalar function of $x$, it corresponds to the independance polynomial~\cite{}.
\begin{equation}
I(G, x) = \sum_{k=1}^{\alpha(G)} a_k x^k,
\end{equation}
where $a_k$ is the the degeneracy for independant set size $k$.

@@

\subsection{Finite field algebra}

>>>>

%\documentclass[a4paper,superscriptaddress,11pt]{quantumarticle}
%\documentclass[aps,twocolumn,longbibliography,english,superscriptaddress]{revtex4-1}
\documentclass{article}
\usepackage{iclr2021_conference}
%\documentclass[a4paper,superscriptaddress,11pt]{article}
\pdfoutput=1
\usepackage[colorlinks=true,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{tabularx}
\usepackage{quoting}
\usepackage{upquote}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage[framemethod=TikZ]{mdframed}
\usepackage{wrapfig}
%\usepackage{caption}
%\usepackage[plain]{algorithm}
\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{rotating}
%\usepackage{cite}
\usepackage{booktabs}
%\usepackage{unicode-math}
%\usepackage{algorithm}% http://ctan.org/pkg/algorithm
%\usepackage{algpseudocode}% http://ctan.org/pkg/algpseudocode
\usepackage{xcolor}% http://ctan.org/pkg/xcolor
\makeatletter
\newsavebox{\@brx}
\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\makeatother
\usepackage{bbm}
\usepackage{jlcode}
\usepackage{graphicx}
\usepackage{amsmath,color,amsthm}
\usepackage{mathrsfs}
\usepackage{float}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{indentfirst}
\usepackage{txfonts}
\usepackage[epsilon, tsrm, altpo]{backnaur}

@@

\begin{abstract}
	Solving the maximum independent set size problem: the maximum independent set size, the independence polynomial and the optimal configuration.
\end{abstract}

@@

For larger graph, we hight recommend using two additional packages,
\texttt{TropicalGEMM} for BLAS speed tropical matrix multiplication and
\texttt{OMEinsumContractionOrders} for hyper optimized contraction order (the KaHyPar~\cite{Schlag2020} approach) ~\cite{Gray2021,Pan2021}.

@@

Then we contract this graph with einsum \texttt{a,b,c,d,ab,ad,bc,bd,cd,ce->}.
The result is a scalar function of $x$, it corresponds to the independence polynomial~\cite{Ferrin2014}.
\begin{equation}
I(G, x) = \sum_{k=1}^{\alpha(G)} a_k x^k,
\end{equation}
where $a_k$ is the the degeneracy for independant set size $k$.

@@

\subsection{Finite field algebra}
It is possible to compute the independence polynomials exactly using basic integer types only,
even when the degeneracy is bigger than that can be represented by 64 bit integers.
What we need is computing with the finite field algebra. In a finite field algebra $GF(p)$, we have the following observations
\begin{enumerate}
    \item One can still use Gaussian elimination~\cite{Golub2013} to solve the linear equations in a finite field, since the multiplicative inverse is properly defined in a finite field,
    \item Let remainders of an integer $x$ over a set of coprime integers $\{p_1, p_2, \ldots, p_n\}$ being $\{x \pmod {p_1}, x\pmod {p_2}, \ldots, x \pmod {p_n}\}$,
    then $\{x \pmod {p_1 \times p_2 \times \ldots \times p_n}\}$ can be obtained using the chinese remainder theorem.
\end{enumerate}
With these observations, we use Algorithm~\ref{alg:finitefield} to compute degeneracies.
In the algorithm, except the computation of chinese remainder theorem, all computations are done with integers with fixed width.

\begin{algorithm}[!ht]
    \small
    \SetAlgoNoLine
    \LinesNumbered
    Let $P = 1$, $X = {0,1,2,\ldots,m}$ and $\hat X_{ij} = X_i^j$, where $i,j = \{0, 1, \ldots m\}$\;
    \While{true}{
        compute the largest prime 64 bit integer $p$ that $\gcd(p, P) = 1$\;
        compute the tensor network contraction on $GF(p)$ and obtain $Y\pmod p = \{y_0 \pmod p, y_1 \pmod p, \ldots , y_m \pmod p\} $\;
        $A \pmod p = \{a_0 \pmod p, a_1 \pmod p, \ldots, a_m \pmod p\} = {\rm gaussian\_elimination}(\hat X, Y \pmod p) $\;
        $A \pmod {P\times p} = {\rm chinese\_remainder}(A \pmod P, A \pmod p)$\;
        \If{$A \pmod P = A \pmod {P \times p}$}{
            \Return $A \pmod P$\;
        }
        $P = P \times p$\;
    }\caption{Compute independence polynomial exactly without integer overflow}\label{alg:finitefield}
\end{algorithm}
